Fast Radio Burst Cosmology

### 2014-06-19 writeup/dm_angular_ps.tex
Minor edits.
Small clean ups.

%\documentclass[onecolumn,amsmath,amssymb,floatfix,12pt,nofootinbib]{revtex4}
\documentclass[twocolumn,prd,noshowpacs,nofootinbib,amsmath,amssymb]{revtex4}
\usepackage{graphicx,epsfig,psfrag,bm,amssymb}
\usepackage{dcolumn}
\usepackage{bm}
\usepackage{mciteplus}
\usepackage{color}
\usepackage{mathrsfs,amsfonts,hepunits, color}
\usepackage{mciteplus}
\usepackage{tikz}
\usepackage{slashed}
\usepackage{verbatim}
\usetikzlibrary{arrows,shapes}
\usetikzlibrary{trees}
\usetikzlibrary{matrix,arrows} 				% For commutative diagram
											% http://www.felixl.de/commu.pdf
\usetikzlibrary{positioning}				% For "above of=" commands
\usetikzlibrary{calc,through}				% For coordinates
\usetikzlibrary{decorations.pathreplacing}  % For curly braces
\usepackage{pgffor}							% For repeating patterns

@@

% for draft comments.
\newcommand{\red}{\textcolor{red}}
\newcommand{\blue}{\textcolor{blue}}
\newcommand{\green}{\textcolor{green}}

@@

To relate observations to cosmology, we wish to determine how a set of observed
dispersion measures $\boldsymbol D$ (whose members are $D_i$)
with known angular positions, $\hat n_i$, but with unknown
redshifts, $z_i$, depends on cosmological parameters.  We begin by defining the
redshift probability distribution function $p(\boldsymbol z)$. We assume that the
redshifts of the FRB sources are independent, \emph{i.e.}~we assume
\begin{equation}
    p(\boldsymbol z) = \prod_i p(z_i).
\end{equation}
This may not be strictly true since FRBs are expected to cluster and trace the
large-scale structure. This is a potential source of bias in the estimator we
derive.  From this assumption it follows that the $D^{\bar{n}_e}_i$ have
uncorrelated probability distributions
\begin{align}
    p(D^{\bar{n}_e}_i)
    &= p(z_i) \left|\frac{\ud D^{\bar{n}_e}}{\ud z} \right|_{z_i}^{-1} \\
        &= p(z_i) \left(\frac{\ud \eta}{\ud z} \right)_{z_i}^{-1},
\end{align}
whose second moment is
\begin{align}
    \langle D^{\bar{n}_e}_i D^{\bar{n}_e}_j \rangle
        &- \langle D^{\bar{n}_e}_i \rangle \langle D^{\bar{n}_e}_j \rangle \\
        &= \delta_{ij} \left[ \int_0^\infty \ud D^{\bar{n}_e} (D^{\bar{n}_e})^2
            p(D^{\bar{n}_e})
            - \left( \int_0^\infty \ud D^{\bar{n}_e} D^{\bar{n}_e}
        p(D^{\bar{n}_e}) \right)^2 \right]\\
        &\equiv \delta_{ij} \sigma_{\bar{n}_e}^2\\
\end{align}

We further assume that $D^{\rm src}_i$ are uncorrelated. Again this may not
be strictly true due to clustering of FRB sources and again this may bias our
estimator. We parameterized the variance from this term as 
\begin{equation}
    \langle D^{\rm src}_i D^{\rm src}_j \rangle
    - \langle D^{\rm src}_i \rangle \langle D^{\rm src}_j \rangle
    = \delta_{ij} \sigma^2_{\rm src}.
\end{equation}

@@

The cosmological information lies in the correlations between the
$D^{\delta_e}_i$. If the redshifts of the observed FRBs where known, we
would write the probability density function of this set as
$p(\boldsymbol D^{\delta_e} | \boldsymbol z)$, which has second moments
\begin{equation}
    w^{z_i z_j}(\hat{n}_i\cdot\hat{n}_j) \equiv \int \ud \boldsymbol D^{\delta_e}
        D^{\delta_e}_i D^{\delta_e}_j p(\boldsymbol D^{\delta_e} | \boldsymbol z).
\end{equation}
$w^{z_i z_j}(\hat{n}_i\cdot\hat{n}_j)$ is the redshift dependant DM angular
correlation function, whose dependence on cosmology is derived below.
Marginalizing $p(\boldsymbol D^{\delta_e} | \boldsymbol z)$ over the unknown redshifts
yields
\begin{equation}
    p(\boldsymbol D^{\delta_e}) = \int_0^\infty \ud \boldsymbol z
        p(\boldsymbol z) p(\boldsymbol D^{\delta_e} | \boldsymbol z).
\end{equation}
The second moment of the marginalized distribution is
\begin{align}
    \langle D^{\delta_e}_i D^{\delta_e}_j \rangle =& \\
        =& \textrm{stuff} \\
        =&
        \begin{cases}
            \int \ud z \ud z' p(z) p(z') w^{z z'}(\hat{n}_i\cdot\hat{n}_j) 
                &  \text{if } (i \neq j) \\
            \int \ud z \ud z' \delta(z - z') p(z) w^{z z'}(\hat{n}_i\cdot\hat{n}_j)
                &  \text{if } (i = j)
        \end{cases}
\end{align}

@@

The errorbar on the redshift averaged dispersion measure angular power spectrum
is
\begin{equation}
    \sigma_{C_l} = \sqrt{\frac{2}{(2 l + 1) f_{\rm sky}}} (C_l + C^N_l),
\end{equation}
where the noise power spectrum is
\begin{equation}
    C^N_l = \frac{4 \pi \sigma_D^2 \sqrt{f_{\rm sky}}}{N_{\rm FRB}}.
\end{equation}
Here $N_{\rm FRB}$ is the number of dispersion measures measured from fast
radio bursts. \red{I'm only 80\% sure about the factors of $f_{\rm sky}$}.

$\sigma_D^2$ is the variance of the dispersion measures.  This would in
practice be measured directly from the observed set, but in terms of
theoretical parameters is
\begin{equation}
\sigma_D^2 = \sigma_{\bar{n}_e}^2 + \sigma^2_{\rm src} + \int \ud z \ud z'
\delta(z - z') p(z) w^{z z'}(1),
\end{equation}
where the first two terms are defined above. The third term is likely
neglidgable and could also be written as
\begin{equation}
    \int \ud z \ud z' \delta(z - z') p(z) w^{z z'}(1) = \int \ud z \ud z'
    \delta(z - z') p(z) \sum_{l} \frac{2l+1}{4\pi} C_l^{zz}.
\end{equation}

>>>>

%\documentclass[onecolumn,amsmath,amssymb,floatfix,12pt,nofootinbib]{revtex4}
\documentclass[onecolumn,prd,noshowpacs,nofootinbib,amsmath,amssymb]{revtex4}
\usepackage{graphicx,epsfig,psfrag,bm,amssymb}
\usepackage{dcolumn}
\usepackage{bm}
\usepackage{mciteplus}
\usepackage{color}
\usepackage{mathrsfs,amsfonts,hepunits, color}
\usepackage{mciteplus}
\usepackage{tikz}
\usepackage{slashed}
\usepackage{verbatim}
\usetikzlibrary{arrows,shapes}
\usetikzlibrary{trees}
\usetikzlibrary{matrix,arrows} 				% For commutative diagram
											% http://www.felixl.de/commu.pdf
\usetikzlibrary{positioning}				% For "above of=" commands
\usetikzlibrary{calc,through}				% For coordinates
\usetikzlibrary{decorations.pathreplacing}  % For curly braces
\usepackage{pgffor}							% For repeating patterns

@@

% for draft comments.
%\newcommand{\red}{\textcolor{red}}
\newcommand{\red}{}
\newcommand{\blue}{\textcolor{blue}}
\newcommand{\green}{\textcolor{green}}

@@

To relate observations to cosmology, we wish to determine how a set of observed
dispersion measures $\boldsymbol D$ (whose members are $D_i$)
with known angular positions, $\hat n_i$, but with unknown
redshifts, $z_i$, depends on cosmological parameters.  We begin by defining the
redshift probability distribution function $p(\boldsymbol z)$. We assume that the
redshifts of the FRB sources are independent, \emph{i.e.}~we assume
\begin{equation}
    p(\boldsymbol z) = \prod_i p(z_i).
\end{equation}
This may not be strictly true since FRBs are expected to cluster and trace the
large-scale structure. This is a potential source of bias in the estimator we
derive.  From this assumption it follows that the $D^{\bar{n}_e}_i$ have
uncorrelated probability distributions
\begin{align}
    p(D^{\bar{n}_e}_i)
    &= p(z_i) \left|\frac{\ud D^{\bar{n}_e}}{\ud z} \right|_{z_i}^{-1} \\
        &= p(z_i) \left(\frac{\ud \eta}{\ud z} \right)_{z_i}^{-1},
\end{align}
whose second moment is
\begin{align}
    \label{e:sig_nebar}
    \langle D^{\bar{n}_e}_i D^{\bar{n}_e}_j \rangle
        &- \langle D^{\bar{n}_e}_i \rangle \langle D^{\bar{n}_e}_j \rangle \\
        &= \delta_{ij} \left[ \int_0^\infty \ud D^{\bar{n}_e} (D^{\bar{n}_e})^2
            p(D^{\bar{n}_e})
            - \left( \int_0^\infty \ud D^{\bar{n}_e} D^{\bar{n}_e}
        p(D^{\bar{n}_e}) \right)^2 \right]\\
        &\equiv \delta_{ij} \sigma_{\bar{n}_e}^2
\end{align}

We further assume that $D^{\rm src}_i$ are uncorrelated. Again this may not
be strictly true due to clustering of FRB sources and again this may bias our
estimator. We parameterized the variance from this term as 
\begin{equation}
    \label{e:sig_src}
    \langle D^{\rm src}_i D^{\rm src}_j \rangle
    - \langle D^{\rm src}_i \rangle \langle D^{\rm src}_j \rangle
    = \delta_{ij} \sigma^2_{\rm src}.
\end{equation}

@@

The cosmological information lies in the correlations between the
$D^{\delta_e}_i$. If the redshifts of the observed FRBs where known, we
would write the probability density function of this set as
$p(\boldsymbol D^{\delta_e} | \boldsymbol z)$, which has second moments
\begin{equation}
    w^{z_i z_j}(\hat{n}_i\cdot\hat{n}_j) \equiv \int \ud \boldsymbol D^{\delta_e}
        D^{\delta_e}_i D^{\delta_e}_j p(\boldsymbol D^{\delta_e} | \boldsymbol z).
\end{equation}
$w^{z_i z_j}(\hat{n}_i\cdot\hat{n}_j)$ is the redshift dependant DM angular
correlation function, whose dependence on cosmology is derived below.
Marginalizing $p(\boldsymbol D^{\delta_e} | \boldsymbol z)$ over the unknown redshifts
yields
\begin{equation}
    p(\boldsymbol D^{\delta_e}) = \int_0^\infty \ud \boldsymbol z
        p(\boldsymbol z) p(\boldsymbol D^{\delta_e} | \boldsymbol z).
\end{equation}
The second moment of the marginalized distribution is
\begin{align}
    \langle D^{\delta_e}_i D^{\delta_e}_j \rangle =& \\
        =&
        \begin{cases}
            \int \ud z \ud z' p(z) p(z') w^{z z'}(\hat{n}_i\cdot\hat{n}_j) 
                &  \text{if } (i \neq j) \\
            \int \ud z \ud z' \delta(z - z') p(z) w^{z z'}(\hat{n}_i\cdot\hat{n}_j)
                &  \text{if } (i = j)
        \end{cases}
\end{align}

@@

The errorbar on the redshift averaged dispersion measure angular power spectrum
is
\begin{equation}
    \sigma_{C_l} = \sqrt{\frac{2}{(2 l + 1) f_{\rm sky}}} (C_l + C^N_l),
\end{equation}
where the noise power spectrum is
\begin{equation}
    C^N_l = \frac{4 \pi \sigma_D^2 \sqrt{f_{\rm sky}}}{N_{\rm FRB}}.
\end{equation}
Here $N_{\rm FRB}$ is the number of dispersion measures measured from fast
radio bursts.
%\red{I'm only 80\% sure about the factors of $f_{\rm sky}$}.

$\sigma_D^2$ is the variance of the dispersion measures.  This would in
practice be measured directly from the observed set, but in terms of
theoretical parameters is
\begin{equation}
\sigma_D^2 = \sigma_{\bar{n}_e}^2 + \sigma^2_{\rm src} + \int \ud z \ud z'
\delta(z - z') p(z) w^{z z'}(1),
\end{equation}
where the first two terms are defined above in equations \ref{e:sig_nebar}
through
\ref{e:sig_src}. The third term is likely
neglidgable and could also be written as
\begin{equation}
    \int \ud z \ud z' \delta(z - z') p(z) w^{z z'}(1) = \int \ud z \ud z'
    \delta(z - z') p(z) \sum_{l} \frac{2l+1}{4\pi} C_l^{zz}.
\end{equation}
